\section{SDG alignment}
We hope this work can also help in the advancement of the Sustainable Development Goals as presented by the United Nations in 2015. In particular, this work aims to be specially helpful in advancing the following goals.

\begin{itemize}
    \item \textbf{(SDG 7) Affordable and clean energy}: By modeling and forecasting solar and wind capacity factor -- and therefore generation -- this work aims to contribute to better integrate renewable energy into the Spanish electrical grid and elsewhere if applicable. Accurate modeling can help in planning infrastructure investments, optimizing resource allocation and reducing fossil fuel reliance, which should drive up affordability and accessibility to clean energy.
    \item \textbf{(SDG 11) Sustainable cities and communities}: Renewable electricity is a key factor in the sustainability transition for cities and communities. Long-term modeling is crucial in ensuring a stable electricity supply for these communities by helping private and public planners make more accurate decisions.
    \item \textbf{(SDG 9) Industry, innovation and infrastructure}: By leveraging some of the latest models in long term time series forecasting, this work aims to bring an innovative approach to modernizing the electrical infrastructure in Spain.
    \item \textbf{(SDG 13) Climate action}: By facilitating the adoption of renewable energy, this work supports the sustainability transition to low carbon energy, directly fighting one of the main drivers of climate change. Long term renewable energy modeling is a very valuable tool for private and public players in reducing greenhouse gas emissions and meeting climate targets.
    \item \textbf{(SDG 8) Decent work and economic growth}: By improving the reliability of the energy infrastructure, this work aims to also help foster economic growth mainly but not only in industrial sectors, which in turn should help create employment opportunities directly in the energy sector and also in all of those helped by having a more robust and affordable electrical grid.
\end{itemize}